---
title: "Analysis of the Impact of the Circle Line and North East Line on Property Prices"
output: html_notebook
runtime: shiny
---

## Introduction
Our project aims to predict the prices of resale HDB flats of various types and the effect that the Circle Line (CCL) and the North East Line (NEL) had on them. We also aim to develop a predictive model that can successfully predict future price fluctuations on the CCL and NEL as well as hopefully the effect of the Downtown Line (DTL) on property prices around it.

## Data Visualization and Modelling

We prepared a linear model with our data on the following variables about each house:

* Distance to the nearest MRT station
* Age of the flat
* Years since the station was built
* Which train station is near it
* What type of flat is it

After we generated the model, we proceeded to plot the visualization below. It is possible to see the actual prices of flats as well our model's predicted flat prices and compare to determine the accuracy of our model.
```{r, echo = FALSE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(sf)
library(Cairo)
options(shiny.usecairo=TRUE)
```

```{r, echo = FALSE, result='hide', warning=FALSE, message=FALSE}
stations <- st_read("~/Documents/DowntownLineAnalysis/shp/CCLNEL Stations Dissolved.shp", quiet=TRUE)
stations.centroid <- st_centroid(stations) %>%
  select(NAME)

ccl_line <- st_read("~/Documents/DowntownLineAnalysis/shp/CCL Rail Line.shp", quiet=TRUE)
nel_line <- st_read("~/Documents/DowntownLineAnalysis/shp/NEL Rail Line.shp", quiet=TRUE)

lines <- c("Circle Line", "North East Line")
completion <- c(2013, 2003)

lineData <- data.frame(lines, completion)

priceShapes <- st_read("~/Documents/DowntownLineAnalysis/shp/CCLNEL Intersection.shp", quiet=TRUE) %>%
  mutate(psm = resale_p_1 / floor_area) %>%
  mutate(year = as.numeric(substring(quarter, 1, 4))) %>%
  filter(year > 1999) %>%
  filter(year < 2018) %>%
  mutate(flat_type = replace(flat_type, flat_type == "MULTI GENERATION", "MULTI-GENERATION")) %>%
  filter(flat_type != "MULTI-GENERATION" & flat_type != "EXECUTIVE") %>%
  mutate(age = year - lease_comm) %>%
  as.data.frame() %>%
  left_join(as.data.frame(stations.centroid), by = c("NAME", "NAME")) %>%
  mutate(unique = paste0(OBJECTID, ".", distance)) %>%
  mutate(unique2 = paste0(block, " ", street_nam)) %>%
  rowwise() %>%
  mutate(actual_distance = st_distance(geometry.x, geometry.y))

for(i in 1:nrow(priceShapes)) {
  priceShapes$years_since_line[i] <- priceShapes$year[i] - lineData$completion[lineData$lines == priceShapes$train.st_2[i]]
}

priceShapes.mins <- priceShapes %>%
  group_by(block, street_nam) %>%
  summarize(min_dist = min(actual_distance)) %>%
  mutate(unique2 = paste0(block, " ", street_nam)) %>%
  left_join(priceShapes, by = c("unique2", "unique2"))
model <- lm(data = priceShapes.mins, formula = psm ~ min_dist + age + years_since_line + factor(train.st_2) + factor(flat_type))
```

```{r, echo = FALSE, results="hide", warning=FALSE, message=FALSE}
subzones <- st_read('~/Documents/DowntownLineAnalysis/shp/Planning_Area_Census2010.shp', quiet=TRUE)
buffers <- st_read("~/Documents/DowntownLineAnalysis/shp/CCLNEL Multiring Buffers.shp", quiet=TRUE) %>%
  mutate(unique = paste0(OBJECTID, ".", distance))
#priceShapes <- st_read("~/Documents/DowntownLineAnalysis/shp/CCLNEL Intersection.shp") %>%
#  mutate(unique = paste0(OBJECTID, ".", distance)) %>%
#  mutate(psm = resale_p_1 / floor_area) %>%
#  mutate(year = as.numeric(substring(quarter, 1, 4))) %>%
#  mutate(flat_type = replace(flat_type, flat_type == "MULTI GENERATION", "MULTI-GENERATION")) %>%
#  filter(flat_type != "MULTI-GENERATION" & flat_type != "EXECUTIVE")
#stations <- st_read("~/Documents/DowntownLineAnalysis/shp/CCLNEL Stations Dissolved.shp")
#stations.centroid <- st_centroid(stations) %>%
#  select(NAME)

colors <- c("#ffffd4", "#fed98e", "#fe9929", "#d95f0e", "#d95f0e")
breaks <- c(0, 0.2, 2, 2.6, 3.2, 1200)
labels <- c("0", "0.2", "2", "2.6", "3.2+", "4+", "bla")
```

```{r, echo = FALSE}
fluidRow(
        column(4, selectInput("line",
                  label = "Choose an MRT line to display",
                  choices = c("Circle Line", "North East Line"),
                  selected = "Circle")),
        column(4, radioButtons("flatType", "Flat Type:",
             choices = sort(unique(priceShapes$flat_type)))),
        column(4, radioButtons("modelType", "Model:",
             choices = c("Actual Prices", "Predicted Prices")))
)
div(class="col-md-12", width="100%", sliderInput(inputId = "year",
            width="100%",
            label = "Year:",
            min = 2000,
            max = max(priceShapes$year),
            value = 2000))
output$unsized <- renderPlot(width=1080, height=1080, {
    line_to_draw <- ccl_line
    color_for_line <- "#DC9F44"
    colors <- c("#ffffd4", "#fed98e", "#fe9929", "#d95f0e", "#d95f0e")
    breaks <- c(0, 0.2, 2, 2.6, 3.2, 1200)
    labels <- c("0", "0.2", "2", "2.6", "3.2+", "4+", "bla")
    
    if(input$line == "North East Line") {
      line_to_draw <- nel_line
      color_for_line <- "#761E86"
      colors <- c("#f2f0f7", "#cbc9e2", "#9e9ac8", "#756bb1", "#54278f")
    }
    if(input$modelType == "Actual Prices") {
      breaks <- c(0, 0.2, 2, 2.6, 3.2, 1200)
      labels <- c("0", "0.2", "2", "2.6", "3.2+", "4+", "bla")
      
      priceShapes.unique <- priceShapes %>%
        filter(train.st_2 == input$line) %>%
        filter(year == input$year) %>%
        filter(flat_type == input$flatType) %>%
        group_by(unique) %>%
        summarise(median_price = median(psm), count = n())
      
      buffers.medians <- buffers %>%
        left_join(as.data.frame(priceShapes.unique), by = c('unique' = 'unique')) %>%
        drop_na(median_price) %>%
        mutate(median_price = median_price / 1000)
      ggplot() +
        geom_sf(subzones, mapping = aes(), fill = "#C6DCEF", color = "#EEEEEE") +
        geom_sf(stations %>% filter(train.st_2 == input$line), mapping = aes(), color = color_for_line, fill = color_for_line) +
        geom_sf(line_to_draw, mapping = aes(), color = color_for_line) +
        xlim(21000, 37000) +
        ylim(26000, 41000) +
        geom_sf(buffers.medians, mapping = aes(fill = cut(median_price,
                                                          include.lowest = T,
                                                          breaks = breaks),
                                               alpha = cut(count,
                                                           breaks = c(0, 5, 10, 10000))), color = NA) +
        coord_sf(datum = NA) +
        scale_fill_manual(values = colors, labels = labels, drop = F, guide = guide_legend(title = "Median House Price",
                                                                title.theme = element_text(size = 25, angle = 0),
                                                                title.position = "top",
                                                                label.position = "top",
                                                                label.theme = element_text(size = 16, angle = 0))) +
        scale_alpha_discrete(guide = "none") +
        theme_bw() +
        theme(legend.position=c(0.8, 0.1), legend.direction = "horizontal",
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              panel.background = element_blank(),
              legend.key.height = unit(8, "mm"),
              legend.key.width = unit(24, "mm"))
    }
    else {
      
      priceShapes.unique <- priceShapes %>%
        filter(train.st_2 == input$line) %>%
        filter(year == input$year) %>%
        filter(flat_type == input$flatType) %>%
        group_by(unique) %>%
        as.data.frame() %>%
        left_join(as.data.frame(stations.centroid), by = c("NAME", "NAME")) %>%
        rowwise()  %>%
        mutate(age = year - lease_comm) %>%
        mutate(years_since_line = year - as.numeric(lineData$completion[lineData$lines == train.st_2])) %>%
        mutate(actual_distance = st_distance(geometry.x, geometry.y)) %>%
        mutate(unique2 = paste0(block, " ", street_nam)) %>%
        ungroup()
      for(i in 1:nrow(priceShapes.unique)) {
        priceShapes.unique$years_since_line[i] <- priceShapes.unique$year[i] - lineData$completion[lineData$lines == priceShapes.unique$train.st_2[i]]
      }
      
      priceShapes.unique <- priceShapes.unique %>%
        group_by(block, street_nam) %>%
        summarize(min_dist = min(actual_distance)) %>%
        mutate(unique2 = paste0(block, " ", street_nam)) %>%
        left_join(priceShapes.unique, by = c("unique2", "unique2"))
      
      priceShapes.unique$f = predict(model, newdata = priceShapes.unique)
      
      priceShapes.unique <- priceShapes.unique %>%
        group_by(unique) %>%
        summarise(median_price = median(f), count = n())
      
      buffers.medians <- buffers %>%
        left_join(priceShapes.unique, by = c('unique' = 'unique')) %>%
        drop_na(median_price) %>%
        mutate(median_price = median_price / 1000)
      ggplot() +
        geom_sf(subzones, mapping = aes(), fill = "#C6DCEF", color = "#EEEEEE") +
        geom_sf(stations %>% filter(train.st_2 == input$line), mapping = aes(), color = color_for_line, fill = color_for_line) +
        geom_sf(line_to_draw, mapping = aes(), color = color_for_line) +
        xlim(21000, 37000) +
        ylim(26000, 41000) +
        geom_sf(buffers.medians, mapping = aes(fill = cut(median_price,
                                                          include.lowest = T,
                                                          breaks = breaks),
                                               alpha = cut(count,
                                                           breaks = c(0, 5, 10, 10000))), color = NA) +
        coord_sf(datum = NA) +
        scale_fill_manual(values = colors, labels = labels, drop = F, guide = guide_legend(title = "Median House Price",
                                                                                           title.theme = element_text(size = 25, angle = 0),
                                                                                           title.position = "top",
                                                                                           label.position = "top",
                                                                                           label.theme = element_text(size = 16, angle = 0))) +
        scale_alpha_discrete(guide = "none") +
        theme_bw() +
        theme(legend.position=c(0.8, 0.1), legend.direction = "horizontal",
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              panel.background = element_blank(),
              legend.key.height = unit(8, "mm"),
              legend.key.width = unit(24, "mm"))
    }
})
renderUI({
  plotOutput("unsized", height = 1080)
})
```

We noticed our model tended to become a lot more accurate in the later years. Upon inspection of the data, we determined it could be due to the much larger amount of data in the later years as opposed to the earlier years. Hence, the model would be better fit for those regions. Out model parameters turned out as follows:
```{r, echo=FALSE}
beautify_summary <- function(model) {

  coefficient_data_frame <- summary(model)$coefficients %>%
    as.data.frame() %>%
    select(Estimate) %>%
    rename(Cofficient = Estimate)
  
  r_squared <- summary(model)$r.squared
  adjusted_r_squared <- summary(model)$adj.r.squared
  
  return(list(paste("R^2 =", r_squared), 
              paste("Adjusted R^2 =", adjusted_r_squared),  
              knitr::kable(coefficient_data_frame)))
}
beautify_summary(model)
beautify_summary(model)[1]
beautify_summary(model)[2]
```

## Predictions

We now attempt to use our model to predict future prices of properties around the CCL and NEL.
```{r, echo = FALSE}
fluidRow(
        column(6, selectInput("line2",
                  label = "Choose an MRT line to display",
                  choices = c("Circle Line", "North East Line"),
                  selected = "Circle")),
        column(6, radioButtons("flatType2", "Flat Type:",
            choices = sort(unique(priceShapes$flat_type))))
)
div(class="col-md-12", width="100%", sliderInput(inputId = "year2",
            width="100%",
            label = "Year:",
            min = 2018,
            max = 2025,
            value = 2018))
# output$unsized2 <- renderPlot(width=1080, height=1080, {
#     line_to_draw2 <- ccl_line
#     color_for_line2 <- "#DC9F44"
#     colors <- c("#ffffe5", "#fff7bc", "#fee391", "#fec44f", "#fe9929", "#ec7014", "#cc4c02", "#8c2d04")
#     breaks <- c(0, 0.2, 2, 2.6, 3.2, 3.3, 3.8, 5, 10000)
#     labels <- c("0", "0.2", "2", "2.6", "3.2", "3.3", "3.8", "4+")
#     if(input$line2 == "North East Line") {
#       line_to_draw2 <- nel_line
#       color_for_line2 <- "#761E86"
#       colors <- c("#fcfbfd", "#efedf5", "#dadaeb", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#4a1486")
#     }
#     priceShapes.unique2 <- priceShapes
#     
#     priceShapes.unique2 <- priceShapes.unique2 %>%
#       filter(train.st_2 == input$line2) %>%
#       mutate(year = input$year2) %>%
#       filter(flat_type == input$flatType2) %>%
#       group_by(unique) %>%
#       as.data.frame() %>%
#       left_join(as.data.frame(stations.centroid), by = c("NAME", "NAME")) %>%
#       rowwise()  %>%
#       mutate(age = year - lease_comm) %>%
#       mutate(years_since_line = year - as.numeric(lineData$completion[lineData$lines == train.st_2])) %>%
#       mutate(actual_distance = st_distance(geometry.x, geometry.y)) %>%
#       mutate(unique2 = paste0(block, " ", street_nam)) %>%
#       ungroup()
#     
#     for(i in 1:nrow(priceShapes.unique2)) {
#       priceShapes.unique2$years_since_line[i] <- priceShapes.unique2$year[i] - lineData$completion[lineData$lines == priceShapes.unique2$train.st_2[i]]
#     }
#     
#     priceShapes.unique2 <- priceShapes.unique2 %>%
#       group_by(block, street_nam) %>%
#       summarize(min_dist = min(actual_distance)) %>%
#       mutate(unique2 = paste0(block, " ", street_nam)) %>%
#       left_join(priceShapes.unique2, by = c("unique2", "unique2"))
#     
#     priceShapes.unique2$f = predict(model, newdata = priceShapes.unique2)
#     
#     priceShapes.unique2 <- priceShapes.unique2 %>%
#       group_by(unique) %>%
#       summarise(median_price = median(f), count = n())
#     
#     buffers.medians <- buffers %>%
#       left_join(priceShapes.unique2, by = c('unique' = 'unique')) %>%
#       drop_na(median_price) %>%
#       mutate(median_price = median_price / 1000)
#     ggplot() +
#       geom_sf(subzones, mapping = aes(), fill = "#C6DCEF", color = "#EEEEEE") +
#       geom_sf(stations %>% filter(train.st_2 == input$line2), mapping = aes(), color = color_for_line2, fill = color_for_line2) +
#       geom_sf(line_to_draw2, mapping = aes(), color = color_for_line2) +
#       xlim(21000, 37000) +
#       ylim(26000, 41000) +
#       geom_sf(buffers.medians, mapping = aes(fill = cut(median_price,
#                                                         include.lowest = T,
#                                                         breaks = breaks)), color = NA) +
#       coord_sf(datum = NA) +
#       scale_fill_manual(values = colors, labels = labels, drop = F, guide = guide_legend(title = "Median House Price",
#                                                                                          title.theme = element_text(size = 25, angle = 0),
#                                                                                          title.position = "top",
#                                                                                          label.position = "top",
#                                                                                          label.theme = element_text(size = 16, angle = 0))) +
#       theme_bw() +
#       theme(legend.position=c(0.7, 0.1), legend.direction = "horizontal",
#             panel.grid.major = element_blank(),
#             panel.grid.minor = element_blank(),
#             panel.border = element_blank(),
#             panel.background = element_blank(),
#             legend.key.height = unit(8, "mm"),
#             legend.key.width = unit(16, "mm"))
# })
output$unsized2 <- renderImage({
  list(src = paste0("~/Documents/DowntownLineAnalysis/plots/", input$line2, "-", input$year2, "-", input$flatType2, ".png"),
         contentType = 'image/png',
         width = 1080,
         height = 1080,
         alt = "This is alternate text")
  }, deleteFile = FALSE)
  
renderUI({
  imageOutput("unsized2", height = 1080)
})
```

