---
title: "Quick and Dirty Analysis of Dataset"
output: html_notebook
---

The goal of this notebook is to perform a quick analysis of the data to see if our hypothesis is explorable or not. We begin by importing necessary libraries and importing data. Our data is from [data.gov](https://data.gov.sg/dataset/resale-flat-prices).

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)

prices <- read_csv('csv/final.csv')
prices
```

Let's see what towns are captured in our dataset.
```{r}
unique(prices$town)
```

We now need to decide on a few towns we wish to analyse over the years. We decided to use the following towns as *"control"* towns:

* Marine Parade
* Hougang
* Woodlands

We chose these due to them being quite far from the Downtown line. These locations also haven't had any new MRT stations nearby in a very long time, which should reduce the possibility of unintended price increases.

We chose the following towns to be our *"experimental"* towns:

* Tampines
* 

These tend to have quite a number of HDB blocks around them, which is what our dataset contains. We now need to extract the relevant components of our dataset and plot facet plots of them.

```{r}
towns <- prices %>%
  filter(town == "MARINE PARADE" | town == "HOUGANG" | town == "WOODLANDS" | town == "TAMPINES") %>%
  mutate(year = substr(month, 1, 4))

ggplot(towns, aes(x=year, fill=median(resale_price))) +
  geom_bar() +
  facet_grid(town ~ ., scales="free_y") +
  theme(strip.text.y = element_text(angle = 0), axis.text.x = element_text(angle = 60, hjust = 1)) +
  ylab("Price in Thousands") + xlab("Year") + 
  ggtitle("Median House Prices") +
  guides(fill=FALSE)
```

